---
title: "Bellabeat Case Study"
author: "mahmoud hassan"
date: "March 2016"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### About a company

Bellabeat, a high-tech manufacturer of health-focused products for women. Bellabeat is a successful small company, but they have the potential to become a larger player in the global smart device market. Urška Sršen, cofounder and Chief Creative Officer of Bellabeat, believes that analyzing smart device fitness data could help unlock new growth opportunities for the company

### Business task
Identify potential opportunities for growth and recommendations for the Bellabeat marketing strategy improvement based on trends in smart device usage.

### Questions for the analysis

* 1-What are some trends in smart device usage?
* 2-How could these trends apply to Bellabeat customers?
* 3-How could these trends help influence Bellabeat marketing strategy



### What Are We Talking About?

How Do **Annual Members** And **Casual Riders** Use Cyclistic bikes Differently? 

* I used Cyclistic’s historical trip data to analyze and identify trends.to visit data source click here [link](https://divvy-tripdata.s3.amazonaws.com/index.html)


##### first install.packages()

```{r}
library("tidyr")
library("tidyverse")
library("lubridate")
library("janitor")
library("dplyr")
library("scales")
library("ggplot2")
```



#### Importing datasets

For this project, I will use FitBit Fitness Tracker [Data](https://www.kaggle.com/datasets/arashnic/fitbit).


#### importing and cleaning data

* Import csv files without blank data which does not read as (NA)
```{r}

setwd("D:/data analysis/capstone project/case study 2/archive (1)/Fitabase Data 4.12.16-5.12.16")
activity <- 
  read.csv("dailyActivity_merged.csv",stringsAsFactors = F)
calories <- 
  read.csv("hourlyCalories_merged.csv",stringsAsFactors =F )
intensities <-
  read.csv("hourlyIntensities_merged.csv",stringsAsFactors =F )
sleep <-
  read.csv("sleepDay_merged.csv",stringsAsFactors = F)
weight <- 
  read.csv("weightLogInfo_merged.csv",stringsAsFactors = F)

```

I already checked the data in Google Sheets. I just need to make sure that everything were imported correctly by using View() and head() functions.

```{r}
head(activity)
```
#### convert data as character to date and time
##### intensities

convert data as character to date and time and add new columns to time and date 

```{r}
intensities$ActivityHour<-as.POSIXct(intensities$ActivityHour,
                          format="%m/%d/%Y %H:%M:%S")
intensities$time<-format(intensities$ActivityHour,format = "%H:%M:%S")
intensities$date<-format(intensities$ActivityHour,"%m:%d:%Y ")

```

##### calories
convert data as character to date and time and add new columns to time and date 

```{r}
calories$ActivityHour<-
      as.POSIXct(calories$ActivityHour, format="%m/%d/%Y %I:%M:%S %p", tz=Sys.timezone())
calories$time<-format(calories$ActivityHour,format = "%H:%M/%S")
calories$date<-format(calories$ActivityHour,format = "%m/%d/%Y")
```

##### weight 

```{r}
weight$Date<-as.POSIXct(weight$Date,format="%m/%d/%Y %I:%M:%S %p", tz=Sys.timezone())
weight$Date<-format(weight$Date,format = "%m/%d/%Y")


```

##### activity

```{r}
activity$ActivityDate<-as.POSIXlt(activity$ActivityDate,format = "%m/%d/%Y")
activity$date <- format(activity$ActivityDate, format = "%m/%d/%y")


```

##### sleep


```{r}
sleep$SleepDay<-as.POSIXlt(sleep$SleepDay,format="%m/%d/%Y %H:%S")
sleep$date <- format(sleep$SleepDay, format = "%m/%d/%y")
```


### Exploring and summarizing data

```{r}
n_distinct(activity$Id)
n_distinct(calories$Id)
n_distinct(intensities$Id)
n_distinct(sleep$Id)
n_distinct(weight$Id)

```
This information tells us about number participants in each data sets.

There is 33 participants in the activity, calories and intensities data sets, 24 in the sleep and only 8 in the weight data set. 8 participants is not significant to make any recommendations and conclusions based on this data.

##### summary statistics of the activity table

```{r}
activity %>% select(TotalSteps,TotalDistance,SedentaryMinutes,Calories,VeryActiveMinutes, FairlyActiveMinutes, LightlyActiveMinutes) %>% 
  summary()
```
* Average sedentary time is 991 minutes or 16 hours. Definately needs to be reduced!
* The majority of the participants are lightly active.





##### summary statistics of the Calories table


```{r}
calories %>% select(Calories) %>% 
  summary()
```

##### summary statistics of the sleep table

```{r}
sleep %>%select(TotalSleepRecords, TotalMinutesAsleep, TotalTimeInBed) %>% 
  summary()

```
* On the average, participants sleep 1 time for 433 Minutes or 7 hours.


##### summary statistics of the weight table

```{r}
weight %>%
  select(WeightKg, BMI) %>%
  summary()
```


#### Merging data
Before beginning to visualize the data, I need to merge two data sets. I’m going to merge (inner join) activity and sleep on columns Id and date (that I previously created after converting data to date time format).

```{r}
merged_data<-merge(activity,sleep,by=c("Id","date"))

head(merged_data)

```

#### Visualization

```{r}

ggplot(data=activity, aes(x=TotalSteps, y=Calories)) + 
  geom_point() + geom_smooth() + labs(title="Total Steps vs. Calories")

```
- I see positive correlation here between Total Steps and Calories, which is obvious -
the more active we are, the more calories we burn.


```{r}

ggplot(data=sleep, aes(x=TotalMinutesAsleep, y=TotalTimeInBed)) + 
  geom_point()+ geom_smooth()+labs(title="Total Minutes Asleep vs. Total Time in Bed")

```
The relationship between Total Minutes Asleep and Total Time in Bed looks linear. So if the Bellabeat users want to improve their sleep, we should consider using notification to go to sleep.



```{r}
int_new <- intensities %>%
  group_by(time) %>%
  drop_na() %>%
  summarise(mean_total_int = mean(TotalIntensity))

ggplot(data=int_new, aes(x=time, y=mean_total_int)) + geom_histogram(stat = "identity", fill='darkblue') +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title="Average Total Intensity vs. Time")

```
* After visualizing Total Intensity hourly, I found out that people are more active between 5 am and 10pm.

* Most activity happens between 5 pm and 7 pm - I suppose, that people go to a gym or for a walk after finishing work. We can **use this time in the Bellabeat app to remind and motivate users to go for a run or walk**.




```{r}

ggplot(data=merged_data, aes(x=TotalMinutesAsleep, y=SedentaryMinutes)) + 
geom_point(color='darkblue') + geom_smooth() +
  labs(title="Minutes Asleep vs. Sedentary Minutes")


```

* Here we can clearly see the negative relationship between Sedentary Minutes and Sleep time.

* As an idea: **if Bellabeat users want to improve their sleep, Bellabeat app can recommend reducing sedentary time**.

* Keep in mind that we need to support this insights with more data, because correlation between some data doesn’t mean causation.

#### recommendations for the business

After analyzing FitBit Fitness Tracker Data, I found some insights that would help influence Bellabeat marketing strategy.

* Women who work full-time jobs (according to the hourly intensity data) and spend a lot of time at the computer/in a meeting/ focused on work they are doing (according to the sedentary time data).

* These women do some light activity to stay healthy (according to the activity type analysis). Even though they need to improve their everyday activity to have health benefits. They might need some knowledge about developing healthy habits or motivation to keep going.








##### The key message for the Bellabeat online campaign

The Bellabeat app is not just another fitness activity app. It’s a guide (a friend) who empowers women to balance full personal and professional life and healthy habits and routines by educating and motivating them through daily app recommendations.

##### Ideas for the Bellabeat app

* As there is no gender information about the participants, I assumed that all genders were presented and balanced in this data set.

* If users want to lose weight, it’s probably a good idea to control daily calorie consumption. Bellabeat can suggest some ideas for low-calorie lunch and dinner.

* If users want to improve their sleep, Bellabeat should consider using app notifications to go to bed.

* Most activity happens between 5 pm and 7 pm - I suppose, that people go to a gym or for a walk after finishing work. Bellabeat can use this time to remind and motivate users to go for a run or walk.

* As an idea: if users want to improve their sleep, the Bellabeat app can recommend reducing sedentary time.

